### 2017-01-16 src/QCAnalysis.Rmd
QC As of now

---
title: "Mamary Gland Development Exploratory Data Analysis"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: united
    highlight: tango
---
***
# Load Data and Packages
```{r, message=FALSE}
library(scran)
library(dplyr)
library(knitr)
library(ggplot2)
rnd_seed <- 300
library(Rtsne)
source("functions.R")
dataList <- readRDS("../data/Robjects/ExpressionList.rds")
m <- dataList[["counts"]]
pD <- dataList[["phenoData"]]
fD <- dataList[["featureData"]]
```
# Overview over the data

## Number of Cells
We sequenced less cells than expected, with a total of `r ncol(m)` cells.
```{r}
kable(t(as.matrix(table(pD$SampleID))))
kable(t(as.matrix(table(pD$Condition))))
```

# Quality Meassures
## Sequencing Depth and Genes detected
```{r}
pD$UmiSums<- colSums(m)
pD$GenesDetected <- colSums(m!=0)
genesDetected <- ggplot(pD, aes(x=SampleID,y=log10(GenesDetected),fill=Condition)) +
    geom_violin(draw_quantiles=0.5)+
    theme_bw()
LibrarySize <- ggplot(pD, aes(x=SampleID,y=log10(UmiSums),fill=Condition)) +
    geom_violin(draw_quantiles=0.5)+
    theme_bw()
gdLibSize <- ggplot(pD, aes(y=log10(GenesDetected), x=log10(UmiSums), color=Condition, shape=Replicate)) +
    geom_point() +
    geom_smooth(method=loess,inherit.aes=FALSE,aes(y=log10(GenesDetected),x=log10(UmiSums))) +
    theme_bw()
genesDetected
LibrarySize
gdLibSize
```

## Genewise QC
```{r, fig.height=8}
ntop <- 50 
mRel <- t(t(m)/colSums(m))
rownames(mRel)  <- fD$symbol
topExpressed <- rowMedians(mRel)
names(topExpressed) <- rownames(mRel)
topExpressed <- topExpressed %>% sort(.,decreasing=TRUE) %>% names
plotData <- t(mRel)[,topExpressed[1:ntop]] %>%
    reshape2::melt() %>%
    rename(Cell=Var1,
	   Gene=Var2,
	   RelativeExpression=value)
topGenes <- ggplot(plotData, aes(x=Gene, y=RelativeExpression)) +
    geom_boxplot() +
    coord_flip() +
    theme_bw()
freqOfExp <- m!=0
rownames(freqOfExp) <- fD$symbol
freqOfExp <- sort(rowSums(freqOfExp)/ncol(freqOfExp),decreasing=TRUE)
plotData <- data.frame("Gene"=names(freqOfExp),"Frequency"=freqOfExp)
topFreq <- ggplot(plotData[1:ntop,], aes(x=factor(Gene,levels=Gene),y=Frequency)) +
    geom_bar(stat="identity") +
    coord_flip() +
    xlab("Gene") +
    theme_bw()
topGenes
topFreq
```

## Cell Viability
```{r}
mMito <- m[fD$Mitochondrial,]
pD$prcntMito <- colSums(mMito)/colSums(m)
cellViability <- ggplot(pD, aes(x=prcntMito, y=GenesDetected, color=Condition, shape=Replicate))+
    geom_point() +
    theme_bw()
cellViability
```
```{r}
smryByGroup <- group_by(pD,Condition) %>%
    summarize(medianPrcntMito=median(prcntMito),
	      madPrcntMito=mad(prcntMito),
	      threshold_PrcntMito=0.05,
	      mGenesDetected=median(log10(GenesDetected)),
	      madGenesDetected=mad(log10(GenesDetected)),
	      threshold_GenesDetected=max(mGenesDetected-3*madGenesDetected,log10(500)),
	      mUmiSums=median(log10(UmiSums)),
	      madUmiSums=mad(log10(UmiSums)),
	      threshold_UmiSums=max(mUmiSums-3*madUmiSums,log10(900))) %>%
    select(Condition,starts_with("threshold")) 
kable(smryByGroup)
```

## QC Thresholding
```{r}
#simple thresholds
MitoCutOff <- 0.05
```
```{r}
#Thresholding per group
#inititae Df
pD <- mutate(pD,
	     ThresholdViability = 0,
	     ThresholdGenesDet = 0,
	     ThresholdLibSize = 0)

grps <- as.character(unique(pD$Condition))
for (grp in grps) {
    thrs <- filter(smryByGroup, Condition==grp) %>% select(-Condition) %>% t() %>% as.vector()
    names(thrs) <- filter(smryByGroup, Condition==grp) %>% select(-Condition) %>% t() %>% rownames()
    pD <- mutate(pD,
		 ThresholdViability= ifelse(Condition==grp, MitoCutOff, ThresholdViability),
		 ThresholdGenesDet= ifelse(Condition==grp, 10^thrs["threshold_GenesDetected"],ThresholdGenesDet),
		 ThresholdLibSize= ifelse(Condition==grp, 10^thrs["threshold_UmiSums"],ThresholdLibSize))
}

pD <- mutate(pD,
	     PassViability=prcntMito < ThresholdViability,
	     PassGenesDet=GenesDetected > ThresholdGenesDet,
	     PassLibSize=UmiSums > ThresholdLibSize,
	     PassAll= PassViability & PassGenesDet & PassLibSize)

gdHist <- ggplot(pD, aes(x=log10(GenesDetected),y=..density..)) +
    geom_histogram(fill="white",color="black") +
    geom_vline(data=smryByGroup,aes(xintercept=threshold_GenesDetected),color="red",lty="longdash") +
    facet_wrap(~Condition) +
    theme_bw()
gdHist

libSizeHist <- ggplot(pD, aes(x=log10(UmiSums),y=..density..)) +
    geom_histogram(fill="white",color="black") +
    geom_vline(data=smryByGroup,aes(xintercept=threshold_UmiSums),color="red",lty="longdash") +
    facet_wrap(~Condition) +
    theme_bw()
libSizeHist

cellViability <- cellViability %+% pD
cellViability <- cellViability + aes(color=PassViability) +
    annotate("rect",ymin=-Inf, ymax=Inf, xmax=Inf, xmin=MitoCutOff
	     ,fill="grey", alpha=0.3) 
cellViability

kable(t(as.matrix(table(pD$Condition,pD$PassGenesDet))),caption="Cells passing Gene Detection QC")
kable(t(as.matrix(table(pD$Condition,pD$PassLibSize))),caption="Cells passing Library Size QC")
kable(t(as.matrix(table(pD$Condition,pD$PassViability))),caption="Cells passing Viability QC")
kable(t(as.matrix(table(pD$Condition,pD$PassAll))),caption="Cells passing all QC requirements")
#Apply Thresholds
pD.filtered <- filter(pD, PassAll)
m.filtered <- m[,as.character(pD.filtered$barcode)]

#Thresholding on lowly expressed genes
isexpThreshold <- 10
expThreshold <- 100*isexpThreshold
keep1 <- rowSums(m.filtered!=0) > isexpThreshold
keep2 <- rowSums(m.filtered) > expThreshold
fD$keep <- keep1 & keep2
m.filtered <- m.filtered[fD$keep,]
fD.filtered <- fD[fD$keep,]
#Final matrix
dim(m.filtered)

qcPCA <- select(pD.filtered, GenesDetected, UmiSums, prcntMito) %>%
    mutate(GenesDetected=log10(GenesDetected),
	   UmiSums=log10(UmiSums))
qcpc <- prcomp(qcPCA)
qcpD <- pD.filtered %>%
    mutate(PC1=qcpc$x[,1],
	   PC2=qcpc$x[,2])
qcpcaPlot <- ggplot(qcpD, aes(x=PC1, y=PC2, color=Condition, shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    ggtitle("PCA plot on QC stats after QC") +
    theme_bw() 
qcpcaPlot

```

## Normalization and first visualization
```{r}
clusters <- quickCluster(m.filtered)
pD.filtered$sf <- computeSumFactors(m.filtered,clusters=clusters,positive=TRUE)
plot(log10(colSums(m.filtered))~log10(pD.filtered$sf),main="Library Size versus Size Factors")
m.norm <- t(t(m.filtered)/pD.filtered$sf)
m.norm <- m.norm[,!is.na(colSums(m.norm))]
pD.filtered <- filter(pD.filtered, barcode %in% colnames(m.norm))
# 
fD.filtered <- mutate(fD.filtered, meanExpression=rowMeans(m.norm),
	     CV2=apply(m.norm,1,cv2),
	     dm=DM(meanExpression,CV2))

highVar <- dplyr::arrange(fD.filtered,desc(dm)) %>% .$id
topX <- floor(0.2*length(highVar))

fD.filtered$highVar <- fD.filtered$id %in% highVar[1:topX]
VarPlot <- ggplot(fD.filtered, aes(x=log10(meanExpression),y=log10(CV2), color=highVar)) +
    geom_point() +
    theme_bw()
VarPlot

fPCA <- log2(t(m.norm[fD.filtered$highVar,])+1)
fPCA <- scale(fPCA,scale=TRUE,center=TRUE)
pc <- prcomp(fPCA)
set.seed(rnd_seed)
tsn <- Rtsne(fPCA,perplexity=floor(nrow(fPCA)/5))

pD.filtered$PC1 <- pc$x[,1]
pD.filtered$PC2 <- pc$x[,2]
pD.filtered$PC3 <- pc$x[,3]
pD.filtered$tSNE1 <- tsn$Y[,1]
pD.filtered$tSNE2 <- tsn$Y[,2]

Acta2 <- filter(fD.filtered, symbol=="Acta2") %>% .$id
pD.filtered$Acta2 <- t(m.norm)[,Acta2]
Fcer1g <- filter(fD.filtered, symbol=="Fcer1g") %>% .$id
pD.filtered$Fcer1g <- t(m.norm)[,Fcer1g]
# 
pcaPlot <- ggplot(pD.filtered, aes(x=PC1, y=PC2, color=Condition, shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw() 
pcaPlot

pcaPlot <- ggplot(pD.filtered, aes(x=PC1, y=PC2, color=log2(Acta2+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw() 
pcaPlot

tsnePlot <- ggplot(pD.filtered, aes(x=tSNE1, y=tSNE2, color=Condition, shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw()
tsnePlot

tsnePlot <- ggplot(pD.filtered, aes(x=tSNE1, y=tSNE2, color=log2(Acta2+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw()
tsnePlot

```
***
## Heatmap on top50
```{r, fig.height=10}
gns <- highVar[1:50]
m.ht <- log2(m.filtered[gns,]+1)
rownames(fD) <- fD$id
rownames(m.ht) <- fD[rownames(m.ht),"symbol"]
colanno <- data.frame("Condition"=pD.filtered$Condition)
rownames(colanno) <- as.character(pD.filtered$barcode)
library(pheatmap)
pheatmap(m.ht, show_colnames=FALSE, annotation_col=colanno, show_rownames=TRUE)

```
## Save Data
```{r}
out <- list()
out[[1]] <- m
out[[2]] <- pD
out[[3]] <- fD
saveRDS(out,file="../data/Robjects/ExpressionList_QC.rds")
```