### 2017-02-01 src/ClusterComparison.Rmd
Added Clustercomparison markdown

--- 
title: "Interactive Markdown to explore Clustering"
author: "Karsten Bach"
date: '`r Sys.Date()`'
runtime: shiny
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: united
    highlight: tango
---
***

```{r echo=FALSE, message=FALSE}
library(scran)
library(dplyr)
library(gtools)
library(knitr)
library(ggplot2)
library(dynamicTreeCut)
library(Rtsne)
library(pheatmap)
source("functions.R")

rnd_seed <- 300
clustDat <- readRDS("../data/Robjects/ClusterComparison.rds") 
bootDat <- readRDS("../data/Robjects/ClusterBootstrap.rds")
dataList <- readRDS("../data/Robjects/ExpressionList_QC.rds")

#Commence
m <- dataList[[1]]
pD <- dataList[[2]]
fD <- dataList[[3]]
#Gene and Cell filtering
m <- m[fD$keep,pD$PassAll]
pD <- pD[pD$PassAll,]
fD <- fD[fD$keep,]
#Normalization
m <- t(t(m)/pD$sf)
#Identification of HVG 
# fD <- mutate(fD, meanExpression=rowMeans(m),
#              CV2= apply(m,1,cv2))
# 
# genes2lbl <- c("Acta2", "Krt18", "Csn2", "Wap","Krt14")
# lblDat <- filter(fD, symbol %in% genes2lbl)
# V3 <- ggplot(arrange(fD,highVar), aes(x=meanExpression,y=CV2, color=highVar)) +
#     geom_label(data=lblDat, aes(x=meanExpression, y=CV2, label=symbol), nudge_y=0.1) +
#     geom_point(data=lblDat, aes(x=meanExpression, y=CV2), pch=1, size=4) +
#     geom_point(alpha=0.8) +
#     scale_x_log10() +
#     scale_y_log10() +
#     theme_bw()
# V3

#Average Linkage appears to best capture the dendrogram

clustDat <- filter(clustDat, Statistic=="Average Silhouette Width")
compPlot <- ggplot(clustDat, aes(x=as.character(DeepSplit), y=Value, color=Linkage,
			    group=Linkage)) +
geom_point() +
geom_path() +
facet_grid(.~Dissimilarity, scales="free_y")
compPlot

#Which Linkage is best
m.sub <- m[fD$highVar,]
# trafM <- log2(m.sub+1)
# dms <- c("pearson","spearman","euclidean")
# res <- NULL
# for (dm in dms) {
# Dissimilarity Measure
#     if(dm!="euclidean") {
#     dis <- as.dist((1-cor(trafM,method=dm)))
#     } else {
#         dis <- dist(t(trafM),method=dm)
#     }
# 
# Hierarchical Clustering
#     htree.average <- hclust(dis, method="average")
#     htree.single <- hclust(dis, method="single")
#     htree.complete <- hclust(dis, method="complete")
#     htree.ward <- hclust(dis, method="ward.D2")
# 
#     ccc.average <- cor(cophenetic(htree.average),dis)
#     ccc.single <- cor(cophenetic(htree.single),dis)
#     ccc.complete <- cor(cophenetic(htree.complete),dis)
#     ccc.ward <- cor(cophenetic(htree.ward),dis)
# 
#     tmp <- data.frame("Linkage"=c("Average",
#                                          "Single",
#                                          "Complete",
#                                          "Ward.D2"),
#                              "Cophenetic Correlation Coefficient"=
#                                  c(ccc.average,ccc.single,
#                                    ccc.complete,ccc.ward),
#                              "Dissimilarity"=dm)
#     res <- rbind(res,tmp)
# }
# kable(res)

inputPanel(
	   selectInput("dm", label = "Dissimilarity",
		       choices= c("euclidean","pearson","spearman"), selected="pearson"),

	   selectInput("ds", label = "DeepSplit",
		       choices= c(0,1,2,3,4), selected=0),

	   selectInput("lk", label = "Linkage",
		       choices=c("average","complete","single","ward.D2"), selected="average")
	  )

#Visualize Dendrograms
clustering <- reactive({
    dynamicCluster(t(m.sub), dm=input$dm,
				 lk=input$lk,
				 ds=as.numeric(input$ds),
				 output="Cluster")
})

renderTable({
    clust <- clustering()
    x <- table(clust$cluster,pD$Condition)
    out <- data.frame("Cluster"=rownames(x),
	       "Virgin"=unname(x[,"V"]),
	       "Pregnant"=unname(x[,"P"]),
	       "Lactation"=unname(x[,"L"]),
	       "Involution"=unname(x[,"I"]))
})

renderPlot({
    clust <- clustering()
    nclust <- length(unique(clust$cluster))
    palette(rainbow(nclust))
    htreeWithBars(clust,pD)
}, width=2000, height=900)

renderPlot({
    clust <- clustering()
    nclust <- length(unique(clust$cluster))
    pD$cluster <- factor(clust$cluster)
    tsnPlot <- ggplot(pD, aes(x=tSNE1, y=tSNE2, color=cluster, shape=Replicate)) +
	geom_point(size=4) +
	scale_color_discrete(rainbow(nclust))+
	theme_bw()
    tsnPlot
}, width=800, height=800)

renderPlot({
    clust <- clustering()
    nclust <- length(unique(clust$cluster))
    palette(rainbow(nclust))
    slct <- paste(input$dm,input$lk,input$ds,sep="_")
    boot <- bootDat[[slct]]$bootresult
    boxplot(t(boot),ylim=c(0,1),col=c(1:nclust))

}, width=800, height=800)

renderPlot({
    library(cluster)
    clust <- clustering()
    nclust <- length(unique(clust$cluster))
    palette(rainbow(nclust))
    sils  <- silhouette(x=clust$cluster,dist=clust$dis)
    plot(sils,col=c(1:length(unique(clust$cluster))), border=NA)

}, width=1000, height=1800)

renderPlot({
    clust <- clustering()
    pD$cluster <- clust$cluster
    nclust <- length(unique(clust$cluster))
    ord <- arrange(pD, cluster) 
    simMat <- asSim(clust$dis)
    simMat.ord <- as.matrix(simMat)[as.character(ord$barcode),
				    as.character(ord$barcode)]
    library(pheatmap)
    annoCol <- data.frame("cluster"=as.factor(ord$cluster),
			  "Condition"=ord$Condition)
    rownames(annoCol) <- as.character(ord$barcode)
    ann_colors <- list("cluster"=rainbow(nclust))
    names(ann_colors$cluster) <- c(min(ord$cluster):max(ord$cluster))
    pheatmap(simMat.ord,
	     cluster_rows=FALSE,
	     cluster_cols=FALSE,
	     show_rownames=FALSE,
	     show_colnames=FALSE,
	     annotation_col=annoCol,
	     annotation_colors=ann_colors)
}, width=1200, height=1200)

inputPanel(
	   selectInput("gene", label = "GeneOfInterest",
		       choices= mixedsort(fD$symbol), selected=c("Krt18","Acta2","Csn2","Cd74")
,
		       multiple=TRUE, selectize=FALSE)
	  )

renderPlot({
    clust <- clustering()
    pD$cluster <- clust$cluster

    genes <- input$gene
	 plotGeneDist(m, pD, fD, genes, colorBy="cluster")
   }, width=800, height=800)

#Compare Cluster stability 
# boot1 <- bootDat[["euclidean_average_0"]]

#Clustering
# clust1 <- dynamicCluster(m.sub, dm="euclidean",
#                          lk="average",
#                          ds=0,
#                          output="Cluster")
# clust2 <- dynamicCluster(m.sub, dm="pearson",
#                          lk="average",
#                          ds=0,
#                          output="Cluster")
# clust3 <- dynamicCluster(m.sub, dm="pearson",
#                          lk="ward.D2",
#                          ds=0,
#                          output="Cluster")

#Looking a different clusters in depth
# 
# fPCA <- t(trafM)
# fPCA <- scale(fPCA,scale=TRUE,center=TRUE)
# set.seed(rnd_seed)
# tsn <- Rtsne(fPCA,perplexity=10,initial_dims=5)
# pD$tSNE1 <- tsn$Y[,1]
# pD$tSNE2 <- tsn$Y[,2]
# tsnePlot <- ggplot(pD, aes(x=tSNE1, y=tSNE2, shape=Condition, color=as.factor(cluster))) +
#     geom_point(alpha=0.8,size=2) +
#     theme_bw()
# tsnePlot
# 
# tsnePlot <- ggplot(pD.filtered, aes(x=tSNE1, y=tSNE2, color=log2(Acta2+1), shape=Replicate)) +
#     geom_point(alpha=0.8,size=2) +
#     theme_bw()
# tsnePlot
```