### 2017-06-14 submissions/resubmission/letter.tex

\documentclass{letter}
\usepackage[a4paper]{geometry}
\begin{document}
\signature{Karsten Bach and Walid Khaled}
\begin{letter}{}
    \opening{Dear Ann,}

Thank you for your rapid and constructive review process. We are grateful for the comments from you and the three reviewers, which highlight the importance of our results for the field of mammary gland biology.

We have revised the manuscript to take into account all of the reviewers' comments as well as recently discovered technical problems of Illumina sequencing.
In particular, we have identified that our presented data was in parts compromised by a sequencing artifact referred to as ``index-swapping''.
Below we have outlined in detail how this has affected our data and that our biological conclusion remain unaffected.
In addition, we have repeated the experiment in order to increase our confidence in the presented results.
From this we have learned that \ldots

We hope that you enjoy reading our revised manuscript and look forward to hearing from you soon.
    \closing{Yours sincerely,}

\end{letter}
\end{document}